---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
```

# Re-analysis of consonance and harmonicity in Friedman et al., (2021) data

<!-- badges: start -->
<!-- badges: end -->

This repository contains a re-analysis of Friedman et al., (2021) data about consonance in Bohlen-Pierce dyads/triads. Our starting point their data (under data folder) and the analysis by Bowling (2021).

## Loading the data

```{r}
library(tidyr)
library(ppcor)
library(ggplot2)

source('read_data.R')
head(as_tibble(df))
# BHM   Bowling Harmonicity model
# HPHM  Harrison Pearce Harmonicity model
# SIM   Huthinson and Knopoff Roughness model (Spectral Interference Model)
```
### Replicate Bowling correlations
```{r}
source('correlational_analysis.R')
```

### Replicate Bowling's plot (B and D)

```{r}
source('figure1.R')
print(fig1)
```

### Impact of controversial trimming

```{r}
source('correlational_analysis_after_trimming_pearson.R')
```

### Analysis of varying thresholds
```{r}
source('varying_thresholds.R')
print(combo)
```

## References

Bowling, D. L. (2021). Harmonicity and roughness in the biology of tonal aesthetics. _Music Perception: An Interdisciplinary Journal, 38 (3)_, 331–334.

Friedman, R. S., Kowalewski, D. A., Vuvan, D. T., & Neill, W. T. (2021). Consonance preferences within an unconventional tuning system. _Music Perception: An Interdisciplinary Journal, 38(3)_, 313–330.

